---
title: "Corn Yield Random Forest: SI Visualizations"
author: "Britta Schumacher"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: yeti
    toc: yes
    toc_float: true
    code_folding: hide
---

# Bring in the data, clean it up, and start visualizing.

Let's first bring in the data, clean it up, and make it spatial!
```{r warning = F, message = F}
library(ggplot2)
library(sp)
library(tidyverse)
library(tmap)
library(RColorBrewer)
library(rgeos)
library(spdplyr)
library(viridis)
library(rgdal)
library(rgeos)
library(ggthemes)
library(ggalt)
library(maps)
library(maptools)
library(grid)
library(biscale)
library(ggcorrplot)
library(ggimage)
library(png)
library(gridExtra)
library(cowplot)
library(wesanderson)
library(ggpubr)
library(egg)

# Bring in data
bioph_preds <- readRDS('./data/data-out/ensemble_bioph_preds.RDS')
farmer_preds <- readRDS('./data/data-out/ensemble_farm_preds.RDS')
ctys <- readRDS("./data/county.RDS")
state <- readRDS("./data/states.RDS")
corn_yield <- readRDS("./data/yieldRF.RDS")

# build frr spatial
frr <- read.csv("./data/FRR_data.csv") 
frr <- frr[,1:2]
colnames(frr)<- c("GEOID", "FRR")
frr$GEOID <- str_pad(frr$GEOID, width = 5, side = "left", pad= "0") 

#Make spatial:
frr_shp <- merge(ctys, frr, by = "GEOID")
spplot(frr_shp, "FRR")

getmode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}

#Aggregate counties to the regional boundaries:
library(maptools)
frr_agg <- unionSpatialPolygons(SpP = frr_shp, IDs = frr_shp$FRR)
frr_final <- aggregate(frr_shp["FRR"], frr_agg, FUN = getmode)
```

## SI Figure 1. Correlation matrix for continuous predictors and corn yield in A) biophysical; and b) farm(er) RF ensembles (see SI Table 1, above for variables, units, and descriptions associated with variable names listed below)
```{r}
corn_yield <- readRDS("./data/yieldRF.RDS")
corn_bioph <- corn_yield %>% select(GEOID,YEAR,YIELD,PERC_IRR,FRR,GDD,BV2,BV18,SDI_CDL_AG,BV9,BV4,TP,ELEVATION,S_PH_H2O,SLOPE,BV8,BV19,T_CEC_SOIL,BV15,T_REF_BULK_DENSITY,T_OC)
corn_bioph <- corn_bioph %>% dplyr::select(-c(GEOID,FRR,YEAR))

# show correlations
corr_yield <- round(cor(corn_bioph[1:18]), 2)
p.mat_yield <- cor_pmat(corn_bioph[c(1:18)])
corrplot_bioph <- ggcorrplot(corr_yield, hc.order = FALSE, p.mat = p.mat_yield, outline.col = "white", type = "lower", insig = "blank",lab = F, ggtheme = ggplot2::theme_minimal, colors = c("#6D9EC1", "white", "#E46726"))

ggsave(filename = "./viz/SIFigure1A.png", plot = corrplot_bioph, dpi = 300, width = 7, height = 7)

corn_farm <- corn_yield %>% 
  select(!GEOID) %>% 
  select(YEAR,YIELD,PERC_IRR,FRR,GDD,BV2,BV18,SDI_CDL_AG,BV9,BV4,TP,ELEVATION,S_PH_H2O,SLOPE,BV8,BV19,T_CEC_SOIL,BV15,T_REF_BULK_DENSITY,T_OC,29:39) %>% 
  select(-c(FRR,YEAR))

# show correlations
corr_yield <- round(cor(corn_farm[1:29]), 2)
p.mat_yield <- cor_pmat(corn_farm[c(1:29)])
corrplot_farm <- ggcorrplot(corr_yield, hc.order = FALSE, p.mat = p.mat_yield, outline.col = "white", type = "lower", insig = "blank",lab = F, ggtheme = ggplot2::theme_minimal, colors = c("#6D9EC1", "white", "#E46726")
                            )

ggsave(filename = "./viz/SIFigure1B.png", plot = corrplot_farm, dpi = 300, width = 7, height = 7)

```

## SI Figure 2. Partial dependence plots of important variables from biophysical and farm(er) ensembles (raw data, not standardized); Partial dependence on: A) precipitation of the warmest quarter (BV18); B) % cultivated area in corn (perc_corn); C) growing degree days (GDD); D) irrigation (PERC_IRR); E) Shannon’s Diversity Index (SDI_CDL_AG); F) mean diurnal range (BV2); G) chemicals ($/acre); H) fertilizers ($/acre); I) government receipts ($/acre); and J) insurance ($/acre).

```{r}
pdp.all <- readRDS("./results/ranger-pdp/pdp-data.RDS")

pdp.all.viz <- ggplot(pdp.all) +   
  geom_line(aes(x = VALUE, y = yhat)) + 
  theme_classic() +
  labs(x = "Raw Value",        
       y = "Predicted Yield (bu/acre)")  +
  facet_wrap(~ VAR, scales = "free", ncol = 3)+  
  theme_minimal() +
  theme(strip.text.x = element_text(size=6)) 

tag_facet(pdp.all.viz, size = 3, vjust = 0.9)

ggsave("./viz/SIFigure2.png", dpi = 400, height = 8, width = 6, units = "in")
```

SI Figure 3. Bivariate choropleth constructed by binning county-level average corn yield (bushels/acre) and percent acres cultivated in corn on agricultural lands from 2008-2018 into thirds; each tercile is then paired and binned into distinct categories. Yellow indicates counties with high average corn yields and an agricultural landscape dominated by corn production, while purple indicates counties with low average yields and a low percentage of agricultural acres in corn. Gray counties indicate missing data.
```{r}
yield_dom <- corn_yield %>% dplyr::select(GEOID, YIELD, YEAR, perc_corn) %>% group_by(GEOID) %>%  summarise(YIELD = mean(YIELD), perc_corn = mean(perc_corn))

#first, build plot to show break-points and number of counties in each bin
#a.v is corn yield = y, p.v. is proportion corn acres = x
y.v<-quantile(yield_dom$YIELD,c(0.33,0.66,1), na.rm = T)
p.v<-quantile(yield_dom$prop_corn,c(0.33,0.66,1), na.rm = T)

yield_dom_new <- yield_dom %>% 
  mutate(y = ifelse(YIELD<y.v[1],1,ifelse(YIELD<y.v[2],2,3)) ,
                     x= ifelse(perc_corn<p.v[1],1,ifelse(perc_corn<p.v[2],2,3))) 

p <- ggplot(data=yield_dom_new,aes(x=perc_corn,y=YIELD,color=atan(y/x),alpha=x+y))+
  geom_point(size=1)+  
  guides(alpha=F,color=F)+
  geom_hline(yintercept=y.v,color="gray20",linetype=2)+
  geom_vline(xintercept=p.v,color="gray20",linetype=2)+
  scale_color_viridis()+
  theme_minimal()+
  theme(plot.caption=element_text(size = 9, hjust=0),
        panel.grid=element_blank()) +
  labs(x="Corn dominance on agricultural land",
       y="Average corn yield (bushels/acre)") 
p

# Bivariate chloropleth (proportion corn acres/ag acres & corn yield) ~ 2017
# subset data and create variables
yield_bivar <- bi_class(yield_dom, x = YIELD, y = perc_corn, style = "quantile", dim = 3)

yield_bivar <- yield_bivar %>% mutate(bi_name = ifelse(bi_class == "1-1", "LYLPC",
ifelse(bi_class == "1-2", "LYMPC",
ifelse(bi_class == "1-3", "LYHPC",
ifelse(bi_class == "2-1", "MYLPC",
ifelse(bi_class == "2-2", "MYMPC",
ifelse(bi_class == "2-3", "MYHPC",
ifelse(bi_class == "3-1", "HYLPC",
ifelse(bi_class == "3-2", "HYMPC",
ifelse(bi_class == "3-3", "HYHPC",
 "NA"))))))))))

yield_bivar$bi_name <- factor(yield_bivar$bi_name, levels = c("LYLPC",
"LYMPC","LYHPC","MYLPC","MYMPC","MYHPC","HYLPC","HYMPC","HYHPC"))

d_complete <- merge(ctys, yield_bivar, by = "GEOID", all = T)
library(tmap)
library(viridisLite)

pal <- viridis(9)

map <- tm_shape(d_complete) +
  tm_polygons("bi_name",
    palette = pal, 
    border.col = "transparent", 
    legend.show = FALSE) +
    tm_layout(frame = FALSE,
            legend.hist.size = 0.5) +
  tm_shape(frr_final) +
  tm_borders(col = "black")

map

tmap_save(map, file = paste0('./viz/SIFigure3-nolegend.png'), dpi = 400)

```

SI Figure 4.

```{r}
# box whisker comparing difficult to regress places and easy v. irrigated extent
ibw <- bioph_preds
ibw$cat <- ifelse(ibw$error < -10 | ibw$error > 10, "Greater than |10| bu/acre", "Less than |10| bu/acre")
ibw$cat <- as.factor(ibw$cat)

ibw <- merge(ibw, corn_yield, by = c("GEOID", "YEAR", "YIELD"))

bw <- ggplot(ibw, aes(x=YEAR, y=PERC_IRR, fill=cat)) +
  geom_boxplot() +
  theme_classic() +
  theme(legend.position = "bottom") +
  theme(axis.title=element_text(size=14),
        axis.text=element_text(size=11)) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_text(size=9, angle = 45, hjust = 1)) +
  ylab("Percent Irrigated Agricultural Land") +
  scale_fill_manual(values=c("#99CCFF", "#4A7DA5")) +
  coord_cartesian(ylim = c(0, 25))  +
  guides(fill=guide_legend(title="Error"))
bw

ggsave(filename = "./viz/SIFigure4.png", plot = bw, dpi = 400, width = 7, height = 5)
```

