---
title: "Corn Yield Random Forest: SI Visualizations"
author: "Britta Schumacher"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: yeti
    toc: yes
    toc_float: true
    code_folding: hide
---

# Bring in the data, clean it up, and start visualizing.

Let's first bring in the data, clean it up, and make it spatial!
```{r warning = F, message = F}
library(ggplot2)
library(sp)
library(tidyverse)
library(tmap)
library(RColorBrewer)
library(rgeos)
library(spdplyr)
library(viridis)
library(rgdal)
library(rgeos)
library(ggthemes)
library(ggalt)
library(maps)
library(maptools)
library(grid)
library(biscale)
library(ggcorrplot)
library(ggimage)
library(png)
library(gridExtra)
library(cowplot)
library(wesanderson)
library(ggpubr)
library(egg)

# Bring in data
bioph_preds <- readRDS('./data/data-out/ensemble_bioph_preds.RDS')
farmer_preds <- readRDS('./data/data-out/ensemble_farm_preds.RDS')
ctys <- readRDS("./data/county.RDS")
state <- readRDS("./data/states.RDS")
corn_yield <- readRDS("./data/yieldRF.RDS")
```

## SI Figure 1. Correlation matrix for continuous predictors and corn yield in A) biophysical; and b) farm(er) RF ensembles (see SI Table 1, above for variables, units, and descriptions associated with variable names listed below)
```{r}
corn_yield <- readRDS("./data/yieldRF.RDS")
corn_bioph <- corn_yield %>% select(GEOID,YEAR,YIELD,PERC_IRR,FRR,GDD,BV2,BV18,SDI_CDL_AG,BV9,BV4,TP,ELEVATION,S_PH_H2O,SLOPE,BV8,BV19,T_CEC_SOIL,BV15,T_REF_BULK_DENSITY,T_OC)
corn_bioph <- corn_bioph %>% dplyr::select(-c(GEOID,FRR,YEAR))

# show correlations
corr_yield <- round(cor(corn_bioph[1:18]), 2)
p.mat_yield <- cor_pmat(corn_bioph[c(1:18)])
corrplot_bioph <- ggcorrplot(corr_yield, hc.order = FALSE, p.mat = p.mat_yield, outline.col = "white", type = "lower", insig = "blank",lab = F, ggtheme = ggplot2::theme_minimal, colors = c("#6D9EC1", "white", "#E46726"))

ggsave(filename = "./viz/SIFigure1A.png", plot = corrplot_bioph, dpi = 300, width = 7, height = 7)

corn_farm <- corn_yield %>% 
  select(!GEOID) %>% 
  select(YEAR,YIELD,PERC_IRR,FRR,GDD,BV2,BV18,SDI_CDL_AG,BV9,BV4,TP,ELEVATION,S_PH_H2O,SLOPE,BV8,BV19,T_CEC_SOIL,BV15,T_REF_BULK_DENSITY,T_OC,29:39) %>% 
  select(-c(FRR,YEAR))

# show correlations
corr_yield <- round(cor(corn_farm[1:29]), 2)
p.mat_yield <- cor_pmat(corn_farm[c(1:29)])
corrplot_farm <- ggcorrplot(corr_yield, hc.order = FALSE, p.mat = p.mat_yield, outline.col = "white", type = "lower", insig = "blank",lab = F, ggtheme = ggplot2::theme_minimal, colors = c("#6D9EC1", "white", "#E46726")
                            )

ggsave(filename = "./viz/SIFigure1B.png", plot = corrplot_farm, dpi = 300, width = 7, height = 7)

```

## SI Figure X. Partial dependence plots of important variables from biophysical and farm(er) ensembles (raw data, not standardized); Partial dependence on: A) precipitation of the warmest quarter (BV18); B) % cultivated area in corn (perc_corn); C) growing degree days (GDD); D) irrigation (PERC_IRR); E) Shannon’s Diversity Index (SDI_CDL_AG); F) mean diurnal range (BV2); G) chemicals ($/acre); H) fertilizers ($/acre); I) government receipts ($/acre); and J) insurance ($/acre).

```{r}
pdp.all <- readRDS("./results/ranger-pdp/pdp-data.RDS")

pdp.all.viz <- ggplot(pdp.all) +   
  geom_line(aes(x = VALUE, y = yhat)) + 
  theme_classic() +
  labs(x = "Raw Value",        
       y = "Predicted Yield (bu/acre)")  +
  facet_wrap(~ VAR, scales = "free", ncol = 3)+  
  theme_minimal() +
  theme(strip.text.x = element_text(size=6)) 

tag_facet(pdp.all.viz, size = 3, vjust = 0.9)

ggsave("./viz/SIFigureX.png", dpi = 400, height = 8, width = 6, units = "in")
```

