---
title: "Tune Corn Yield Random Forest: Biophysical & Farm(er)"
author: "Britta Schumacher"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: yeti
    toc: yes
    toc_float: true
    code_folding: hide
---

# Bring in the data, clean it up, and run the model on the training set
```{r packages and data}
library(randomForest)
library(tidyverse)
library(ggplot2)
library(caret)
library(ie2misc)
library(MLmetrics)

# pull in TRUE train data; all data with recorded yield
corn_yield <- readRDS("./data/yieldRF.RDS")

# remove ID variables that aren't included in the model
yield_RF <- corn_yield %>% dplyr::select(-c(GEOID))

# specify qualitative variables that are factors
yield_RF$YEAR <- as.factor(yield_RF$YEAR)
yield_RF$FRR <- as.factor(yield_RF$FRR)

```

# Build training and test sets for: all data, pruned biophysical data, and pruned farm(er) data
```{r train and test}
set.seed(1234) 
random_rn <- sample(nrow(yield_RF), ceiling(nrow(yield_RF)*.25)) 
train_fullset <- yield_RF[-random_rn,] 
test_fullset <- yield_RF[random_rn,] 

# biophysical
train <- train_fullset %>% select(YEAR,YIELD,PERC_IRR,FRR,GDD,BV2,BV18,SDI_CDL_AG,BV9,BV4,TP,ELEVATION,S_PH_H2O,SLOPE,BV8,BV19,T_CEC_SOIL,BV15,T_REF_BULK_DENSITY,T_OC)
test <- test_fullset %>% select(YEAR,YIELD,PERC_IRR,FRR,GDD,BV2,BV18,SDI_CDL_AG,BV9,BV4,TP,ELEVATION,S_PH_H2O,SLOPE,BV8,BV19,T_CEC_SOIL,BV15,T_REF_BULK_DENSITY,T_OC)

# farmer
farmer_train <- train_fullset %>% select(YEAR,YIELD,PERC_IRR,FRR,GDD,BV2,BV18,SDI_CDL_AG,BV9,BV4,TP,ELEVATION,S_PH_H2O,SLOPE,BV8,BV19,T_CEC_SOIL,BV15,T_REF_BULK_DENSITY,T_OC,29:39)
farmer_test <- test_fullset %>% select(YEAR,YIELD,PERC_IRR,FRR,GDD,BV2,BV18,SDI_CDL_AG,BV9,BV4,TP,ELEVATION,S_PH_H2O,SLOPE,BV8,BV19,T_CEC_SOIL,BV15,T_REF_BULK_DENSITY,T_OC,29:39)
```

# Tune Biophysical RF
```{r tune biophysical RF, evaluate = FALSE}
control <- trainControl(method="repeatedcv", 
                        number=10, 
                        repeats=3, 
                        search="grid")
set.seed(1234)

tunegrid <- expand.grid(.mtry=c(1:12))

rf_gridsearch <- train(YIELD~., data=train, 
                       method="rf", 
                       metric='RMSE', 
                       ntree = 2000,
                       tuneGrid=tunegrid,
                       trControl = control)
print(rf_gridsearch)

jpeg("./viz/rf-tune-mtry2.jpeg", height = 4, width = 6, units = "in", res = 400)
plot(rf_gridsearch)
dev.off()
```

Random Forest 

12737 samples
   19 predictor ~ default ntrees = 500

No pre-processing
Resampling: Cross-Validated (10 fold, repeated 3 times) 
Summary of sample sizes: 11463, 11463, 11464, 11463, 11462, 11464, ... 
Resampling results across tuning parameters:

  mtry  RMSE      Rsquared   MAE     
   1    30.08711  0.6442456  23.84962
   2    21.61608  0.7622326  16.78247
   3    19.18015  0.7934285  14.70176
   4    18.48832  0.8016640  14.10895
   5    18.22066  0.8042397  13.87249
   6    18.10367  0.8048134  13.76500
   7    18.02260  0.8051778  13.68915
   8    17.97065  0.8051751  13.64230
   9    17.93063  0.8052253  13.60744
  10    17.93072  0.8044584  13.59464
  11    17.91163  0.8043202  13.57528
  12    17.92377  0.8034832  13.57159

RMSE was used to select the optimal model using the smallest value.
The final value used for the model was mtry = 11.

Random Forest 

12737 samples
   19 predictor ~ ntrees = 2000

No pre-processing
Resampling: Cross-Validated (10 fold, repeated 3 times) 
Summary of sample sizes: 11463, 11463, 11464, 11463, 11462, 11464, ... 
Resampling results across tuning parameters:

  mtry  RMSE      Rsquared   MAE     
   1    30.09296  0.6454033  23.85180
   2    21.59564  0.7636618  16.77102
   3    19.16431  0.7942486  14.69318
   4    18.47087  0.8022988  14.09545
   5    18.20506  0.8048216  13.85975
   6    18.07693  0.8055895  13.74367
   7    18.00142  0.8058497  13.67102
   8    17.94524  0.8059420  13.61931
   9    17.92014  0.8055923  13.59750
  10    17.90729  0.8051589  13.57500
  11    17.89944  0.8046973  13.56337
  12    17.89510  0.8042720  13.55620

RMSE was used to select the optimal model using the smallest value.
The final value used for the model was mtry = 12.

# Biophysical RF variable importance with various mtry levels
```{r biophysical RF variable importance, evaluate = FALSE}
# remove variables that contributed least to the full model
rf_mtry <- function(mtry) {
  
  #index
  index <- mtry
  
  # build RF on training data
  set.seed(1234) 
  rf_yield <- randomForest(YIELD ~ ., data = train, mtry = index, ntree = 2000)
  rf_yield
  
  # build predictions on test set
  preds <- predict(rf_yield, test, type = "response")
  test$PREDICTIONS <- preds
  
  # build MSE/RMSE
  RMSE <- caret::RMSE(test$PREDICTIONS, test$YIELD)
  R2 <- 1 - (sum((test$YIELD - test$PREDICTIONS)^2)/sum((test$YIELD - mean(test$YIELD))^2))
  
  # build clean dataframe
  colnames(test)[2] <- paste0('YIELD_',index)
  colnames(test)[21] <- paste0('PREDS_',index)
  
  # importance plots
  imp <- as.data.frame(importance(rf_yield))
  imp$varnames <- rownames(imp) # row names to column
  
  colnames(imp)[1] <- paste0('IncPurity_',index)
  
  # build argument to feed save, below
  list <- list("RMSE" = RMSE, "R2" = R2, "var_exp" = rf_yield)
  
  # return
  saveRDS(imp, file = paste0('./results-imp/RFimp_',index,'.RDS'))
  return(list)

}

rf_mtry_1 <- rf_mtry(1)
rf_mtry_1
rf_mtry_2 <- rf_mtry(2)
rf_mtry_2
rf_mtry_3 <- rf_mtry(3)
rf_mtry_3
rf_mtry_4 <- rf_mtry(4)
rf_mtry_4
rf_mtry_5 <- rf_mtry(5)
rf_mtry_5
rf_mtry_6 <- rf_mtry(6)
rf_mtry_6
rf_mtry_7 <- rf_mtry(7)
rf_mtry_7
rf_mtry_8 <- rf_mtry(8)
rf_mtry_8
rf_mtry_9 <- rf_mtry(9)
rf_mtry_9
rf_mtry_10 <- rf_mtry(10)
rf_mtry_10
rf_mtry_11 <- rf_mtry(11)
rf_mtry_11
rf_mtry_12 <- rf_mtry(12)
rf_mtry_12
```

# Tune Farm(er) RF
```{r tune farmer RF}
control <- trainControl(method="repeatedcv", 
                        number=10, 
                        repeats=3, 
                        search="grid")
set.seed(1234)

tunegrid <- expand.grid(.mtry=c(5:25))

rf_farmer_gridsearch <- train(YIELD~., data=farmer_train, 
                       method="rf", 
                       metric='RMSE', 
                       ntree = 2000,
                       tuneGrid=tunegrid,
                       trControl = control)
print(rf_farmer_gridsearch)

jpeg("./viz/rf-farmer-tune-mtry.jpeg", height = 4, width = 6, units = "in", res = 400)
plot(rf_farmer_gridsearch)
dev.off()
```
